# (PART) MACHINE LEARNING {-} 

# Machine Learning and AI

## Term: Classification

**Definition:** Classification is a machine learning task where the goal is to predict the categorical class labels of new instances based on past observations or training data.

### Random Forest Model

**R Code Illustration:** (Note: This example demonstrates how to perform classification using the caret package.)

```{r}
# Load required packages
library(caret)

# Load example dataset
data(iris)

# Train a classification model (e.g., Random Forest)
model <- train(Species ~ ., data = iris, method = "rf")

# Print the trained model
cat(model)

```


## Term: Regression
**Definition:** Regression is a machine learning task where the goal is to predict continuous numerical values based on input features, aiming to model the relationship between variables.

### Linear Regression Model

**R Code Illustration:** (Note: This example demonstrates how to perform regression using the caret package.)

```{r}
# Train a regression model (e.g., Linear Regression)
model <- train(Sepal.Length ~ ., data = iris, method = "lm")

# Print the trained model
cat(model)

```


## Term: Feature Selection
**Definition:** Feature selection is the process of selecting a subset of relevant features (input variables) from the original set of features, aiming to improve model performance, reduce overfitting, and enhance interpretability.


### Recursive Feature Elimination

**R Code Illustration:** (Note: This example demonstrates how to perform feature selection using the caret package.)

```{r}
# Perform feature selection (e.g., Recursive Feature Elimination)
selected_features <- rfe(iris[, -5], iris[, 5], sizes = c(1:4), rfeControl = rfeControl(functions = lmFuncs))

# Print selected features
cat(selected_features)

```